Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2013-01-10 1.tex
immediate pre-defense writing

There is only one concise explanation for the [dʒ]-[tʃ] and [k]-[ɡ] alternations, namely that Russian speakers have internalized a process of voice assimilation.
It is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers. 
For instance, \citeauthor{PE} write:

\begin{quote}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their langauge, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for. \citep[17f.]{PE}
\end{quote}

@@

There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

>>>>

There is only one concise explanation for the [dʒ]-[tʃ] and [k]-[ɡ] alternations, namely that Russian speakers have internalized a process of voice assimilation.
It is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers, as some linguists have done.

\begin{quote}
MSCs are merely artifacts of the grammar, and thus play no part in the phonological component of a language. \citep[302]{Clayton1976}

Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for. \citep[17f.]{PE}

[MSCs] have no observable consequences in the course of the normal use of the language\ldots{} \citep[320]{Zimmer1969}
\end{quote}

@@

%FIXME

\begin{example}[Adaptation of onset clusters in Finnish loanwords]
\begin{tabular}{l llll}
a. & toːri    & `store'   & ($<$ \emph{store})              & \citep[English:][89]{Hellstrom1976}   \\
   & rosseri  & `grocery' & ($<$ \emph{grocery})            &                             \\
b. & ranska   & `French'  & ($<$ \emph{Franska})            & \citep[Swedish:][67]{Campbell2004}    \\
   & ruːvi    & `screw'   & ($<$ \emph{skruv})              &                             \\
c. & risti    & `cross'   & ($<$ \emph{kristĭ})             & \citep[Old Russian:][60]{Bjornflaten2006} \\
   & raːmattu & `bible'   & ($<$ \emph{gramota} `document') &                             \\
\end{tabular}
\end{example}

There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt{Fries1949}, \citealt{Holden1976}, \citealt{Ito1995a,Ito1995b}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
For instance, native words in San Mateo Huave all end in a consonant, but final unstressed syllables in Spanish loanwords are never repaired by epenthesis (e.g., \emph{verde} `green' $>$ [beɾde], *[beɾdej]).
%FIXME example here
Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

### 2013-01-10 3.tex
immediate pre-defense writing

It has long been suggested that statistical criteria could distinguish between \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, and those gaps which are \emph{structural} in nature \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
Following seminal work by \citet{Mester1988} and \citet{McCarthy1988}, it is now commonly assumed that the phonotactic patterns in lexical entries ``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004} acquired by speakers, and therefore phonotactic constraints, whether categorical or gradient, can be directly inferred from statistical analysis of the lexicon.
There is by now an enormous amount of research that adopts this assumption.
Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi, \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment. 
\citet{Pozdniakov2007} analyze lexical tendencies in a diverse sample of 30 languages.

>>>>

It has long been suggested that statistical criteria could distinguish between \emph{accidental phonotactic gaps}, those which could arise without any antecedent cause, and those gaps which are \emph{structural} in nature \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}.
Following seminal work by \citet{Mester1988} and \citet{McCarthy1988}, it is now commonly assumed that the phonotactic patterns in lexical entries ``directly determine the mental representation of the phonotactic constraints'' \citep[180]{Frisch2004} acquired by speakers, and therefore phonotactic constraints, whether categorical or gradient, can be directly inferred from statistical analysis of the lexicon.
There is by now an enormous amount of research that adopts this assumption.
Semitic languages, especially Arabic \citep{Coetzee2008a,Frisch2004,McCarthy1988,Pierrehumbert1993} but also Berber \citep{Elmedlaoui1995} and Tigrinya \citep{Buckley1997}, and Austronesian languages, including Javanese \citep{GraffInPress,Mester1988}, Muna \citep{Anttila2008a,Coetzee2008a}, and Samoan \citep{AldereteInPress}, have been of particular interest to linguists adopting this hypothesis; Aymara \citep{GraffInPress}, Cantonese \citep{Yip1989}, English \citep{Berkley1994a,Berkley1994b,Berkley2000,Coetzee2008b,Davis1989,Martin2007,Martin2011}, Gitksan \citep{Brown2010}, Hungarian \citep{Grimes2010}, Japanese \citep{Kawahara2006}, Ju|'hoansi \citep{Kinney2005}, Navajo \citep{Martin2007,Martin2011}, Ofo \citep[38f.]{MacEachern1999}, Russian \citep{Padgett1991,Padgett1992}, Shona \citep{Hayes2008a}, and Wargamay \citep{Hayes2008a} have also received a statistical treatment. 
\citet{Pozdniakov2007} analyze lexical tendencies in a diverse sample of 30 languages.
